### 2017-05-24 authorlist.tex
Pushing overleaf paper

\author{J.~Selsing\inst{1}
\and T. Kr\"{u}hler\inst{2}
\and D.~Malesani\inst{1,3}
\and P. D'Avanzo\inst{4}
\and J.~Palmerio\inst{5}
\and S.~D.~Vergani\inst{6}
\and J.~Japelj\inst{6}
\and B.~Milvang-Jensen\inst{1}
\and D.~Watson\inst{1}
\and P.~Jakobsson\inst{7}
%
\and Z. Cano\inst{8}
\and J.~P.~U.~Fynbo\inst{1}
\and A. Gomboc\inst{9}
\and K.~E.~Heintz\inst{7,1}
\and M. Sparre\inst{1,10}
\and XSGRB collaboration
}

\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
%2
\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
%3
\and DTU Space, National Space Institute, Technical University of Denmark, Elektrovej 327, DK-2800 Lyngby, Denmark
%4
\and INAF - Osservatorio Astronomico di Brera, via E. Bianchi 46, I-23807, Merate (LC), Italy
%5
\and Sorbonne Universités, UPMC Univ. Paris 6 et CNRS, UMR 7095, Institut d’Astrophysique de Paris, 98 bis bd Arago, 75014 Paris, France
%6
\and GEPI, Observatoire de Paris, PSL Research University, CNRS, Univ. Paris Diderot, Sorbonne Paris Cité, 5 Place Jules Janssen, 92195 Meudon, France
%7
\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, 107 Reykjav\'ik, Iceland
%8
\and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain.
%9
\and University of Nova Gorica, Vipavska 13, 5000 Nova Gorica, Slovenia.
%10
\and Heidelberger Institut f{\"u}r Theoretische Studien, Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany 
%\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
}