### 2017-06-26 authorlist.tex

\author{J.~Selsing\inst{1}
\and T.~Kr\"{u}hler\inst{2}
\and D.~Malesani\inst{1,3}
\and P.~D'Avanzo\inst{4}
\and S.~Schulze \inst{5, 6, 7}
\and J.~Palmerio\inst{8}
\and S.~D.~Vergani\inst{9}
\and J.~Japelj\inst{10}
\and B.~Milvang-Jensen\inst{1}
\and D.~Watson\inst{1}
\and P.~Jakobsson\inst{11}
%
\and J.~Bolmer \inst{2}
\and Z.~Cano\inst{12}
\and S.~Covino \inst{13}
%\and L.~B.~Christensen\inst{1}
\and V.~D'Elia\inst{13}
\and A.~de~Ugarte~Postigo\inst{12, 1}
\and J.~P.~U.~Fynbo\inst{1}
\and A.~Gomboc\inst{14}
\and K.~E.~Heintz\inst{11,1}
\and A.~J.~Levan \inst{15}
\and S.~Piranomonte \inst{16}
\and V.~Pugliese \inst{10}
\and R.~S\'{a}nchez-Ram\'{\i}rez \inst{12, 13}
\and M.~Sparre\inst{1,17}
\and N.~R.~Tanvir\inst{18}
\and C.~C.~Th\"one\inst{9}
\and K.~Wiersema \inst{18}
}

\institute{Dark Cosmology Centre, Niels Bohr Institute, University of Copenhagen, Juliane Maries Vej 30, 2100 K\o benhavn \O, Denmark
%2
\and Max-Planck-Institut f\"{u}r extraterrestrische Physik, Giessenbachstra\ss e, 85748 Garching, Germany
%3
\and DTU Space, National Space Institute, Technical University of Denmark, Elektrovej 327, DK-2800 Lyngby, Denmark
%4
\and INAF - Osservatorio Astronomico di Brera, via E. Bianchi 46, I-23807, Merate (LC), Italy
%5
\and Instituto de Astrof\'isica, Facultad de F\'isica, Pontificia Universidad Cat\'olica de Chile, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
%6
\and Millennium Institute of Astrophysics, Vicu\~{n}a Mackenna 4860, 7820436 Macul, Santiago, Chile
%7
\and Department of Particle Physics and Astrophysics, Faculty of Physics, Weizmann Institute of Science, Rehovot 76100, Israel
%8
\and Sorbonne Universités, UPMC Univ. Paris 6 et CNRS, UMR 7095, Institut d’Astrophysique de Paris, 98 bis bd Arago, 75014 Paris, France
%9
\and GEPI, Observatoire de Paris, PSL Research University, CNRS, Univ. Paris Diderot, Sorbonne Paris Cité, 5 Place Jules Janssen, 92195 Meudon, France
%10
\and Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands
%11
\and Centre for Astrophysics and Cosmology, Science Institute, University of Iceland, Dunhagi 5, 107 Reykjav\'ik, Iceland
%12
\and Instituto de Astrof\'isica de Andaluc\'ia (IAA-CSIC), Glorieta de la Astronom\'ia s/n, E-18008, Granada, Spain.
%13
\and INAF-Osservatorio Astronomico di Roma, Via Frascati 33, I-00040 Monteporzio Catone, Italy;
ASI-Science Data Centre, Via del Politecnico snc, I-00133 Rome, Italy
%14
\and Centre for Astrophysics and Cosmology, University of Nova Gorica, Vipavska 13, 5000 Nova Gorica, Slovenia.
%15
\and Department of Physics, University of Warwick, Coventry CV4 7AL, UK
%16
\and INAF, Osservatorio Astronomico di Brera, Via E. Bianchi 46, I-23807 Merate (LC), Italy
%17
\and Heidelberger Institut f{\"u}r Theoretische Studien, Schloss-Wolfsbrunnenweg 35, 69118 Heidelberg, Germany 
%18
\and Department of Physics and Astronomy, University of Leicester, University Road, Leicester, LE1 7RH, UK
%\and European Southern Observatory, Alonso de C\'{o}rdova 3107, Vitacura, Casilla 19001, Santiago 19, Chile
}