### 2017-11-25 tables/spec_overview.tex

\begin{table*}[!ht]

	\centering
	\caption{Overview of the spectroscopic observations. ``JH'' in the slit width refers to observations where a $K$-band blocking filter has been used. The seeing is determined from the width of the spectral trace of a telluric standard star, observed close in time to the host observation. The spectral resolution, $R$, is measured from unresolved telluric absorption lines in the spectrum of the telluric standard star. \label{tab:spec_overview}}
	\begin{tabular}{cccccccc}
		\hline\hline
		{Observation epoch} &  \multicolumn{3}{c}{Exposure time (s)} & Slit width & Airmass & Seeing & $R$   \\ [1.5pt]
        \hline
		(UT) & UVB  & VIS & NIR &  (arcsec)   & {} & (arcsec)  & {VIS/NIR}  \\ [1.5pt]
		\hline
		2011-11-19 01:33 & 2 $\times$ 2400 & 2 $\times$ 2400 & 8 $\times$ 600 & 1.0/1.0/0.9 & 1.49 & 0.75 & 11600/6700 \\
        2013-07-15 09:02 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.53 & 0.98 & 9600/8900 \\
        2013-08-03 07:37 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.55 & 0.85 & 11400/11300 \\
        2013-08-03 08:34 & 2 $\times$ 1200 & 2 $\times$ 1200 & 8 $\times$ 300 & 1.0/1.0/0.9JH & 1.49 & 0.85 & 11400/11300 \\

		\hline\noalign{\smallskip}

	\end{tabular}

\end{table*}